Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-05-06 document/ms.tex
actually mention the importance sampling approximation

In other words, a catalog is a dimensionality reduction of the raw data with
all the relevant information retained.
In the context of \kepler, the catalog reduces the set of downloaded pixel
time series (approximately XXXX data points for the average \kepler\ target)
to probabilistic constraints on a handful of physical parameters---\entry\
from above---like the orbital period and radius.
If we take this set of parameters as \emph{sufficient statistics} of the data
then we can, in theory, compute \eq{crazylike}---up to an unimportant
constant---without ever looking at the raw data again.
This is true because the catalog provides constraints on the posterior
probability of the parameters $\{\entry_k\}$ under some choice of ``interim
prior'' \interim
\begin{eqnarray}\eqlabel{crazypost}
p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
\frac{p(\{\data_k\}\,|\,\{\entry_k\})\,p(\{\entry_k\}\,|\,\interim)}
     {p(\{\data_k\}\,|\,\interim)} \quad.
\end{eqnarray}
It turns out that we can use this posterior to simplify \eq{crazylike} to a
form that can, in many common cases, be evaluated efficiently.
To find this result, multiply the integrand in \eq{crazylike} by
\begin{eqnarray}
\frac{p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
\end{eqnarray}
and use \eq{crazypost} to find
\begin{eqnarray}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &=&
    \int
    \frac{p(\{\entry_k\}\,|\,\ratepars)}{p(\{\entry_k\}\,|\,\interim)}\,
    p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)
    \dd\{\entry_k\} \quad.
\end{eqnarray}
The data only enter this equation through the posterior constraints provided
by the catalog!

A very simple example is a procedure that will be very familiar: the histogram.
If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
constant rate function---where the bin heights are the parameters---and if the
uncertainties on the catalog are negligible compared to the bin widths then
the maximum marginalized likelihood solution for \ratepars\ is a histogram of
the catalog entries (see \app{histogram} for the derivation).
The case of non-negligible uncertainties is described by \citet{hogge} using a
method similar to the one discussed below.

@@

\paragraph{Occurence rate parameterization \& priors}
\dfmtodo{describe the Gaussian process prior on the bin heights.}

>>>>

In other words, a catalog is a dimensionality reduction of the raw data with
all the relevant information retained.
In the context of \kepler, the catalog reduces the set of downloaded pixel
time series (approximately XXXX data points for the average \kepler\ target)
to probabilistic constraints on a handful of physical parameters---\entry\
from above---like the orbital period and radius.
If we take this set of parameters as \emph{sufficient statistics} of the data
then we can, in theory, compute \eq{crazylike}---up to an unimportant
constant---without ever looking at the raw data again.
This is true because the catalog provides constraints on the posterior
probability of the parameters $\{\entry_k\}$ under some choice of ``interim
prior'' \interim
\begin{eqnarray}\eqlabel{crazypost}
p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) &=&
\frac{p(\{\data_k\}\,|\,\{\entry_k\})\,p(\{\entry_k\}\,|\,\interim)}
     {p(\{\data_k\}\,|\,\interim)} \quad.
\end{eqnarray}
It turns out that we can use this posterior to simplify \eq{crazylike} to a
form that can, in many common cases, be evaluated efficiently.
To find this result, multiply the integrand in \eq{crazylike} by
\begin{eqnarray}
\frac{p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
     {p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)}
\end{eqnarray}
and use \eq{crazypost} to find
\begin{eqnarray}\eqlabel{simplemarglike}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &=&
    \int
    \frac{p(\{\entry_k\}\,|\,\ratepars)}{p(\{\entry_k\}\,|\,\interim)}\,
    p(\{\entry_k\}\,|\,\{\data_k\},\,\interim)
    \dd\{\entry_k\} \quad.
\end{eqnarray}
The data only enter this equation through the posterior constraints provided
by the catalog!

The constraints in \eq{crazypost} can always be---and often are---propagated
as a list of $N$ samples $\{\entry_k^{(n)}\}$ from the posterior
\begin{eqnarray}
\entry_k^{(n)} &\sim& p(\{\entry_k\}\,|\,\{\data_k\},\,\interim) \quad.
\end{eqnarray}
We can use these samples to approximately compute the integral in
\eq{simplemarglike} as
\begin{eqnarray}
\frac{p(\{\data_k\}\,|\,\ratepars)}{p(\{\data_k\}\,|\,\interim)} &\approx&
    \frac{1}{N} \sum_{n=1}^N
    \frac{p(\{\entry_k^{(n)}\}\,|\,\ratepars)}
         {p(\{\entry_k^{(n)}\}\,|\,\interim)} \quad.
\end{eqnarray}
This is very efficient to compute as long as an evaluation of
$p(\{\entry_k^{(n)}\}\,|\,\ratepars)$ is not very expensive.
That being said, this could be a high variance estimator depending on the
initial choice of $p(\{\entry_k^{(n)}\}\,|\,\interim)$ but in our case, we'll
show that not many samples are needed to get an accurate estimate.

A very simple example is a procedure that will be very familiar: the histogram.
If you model the function $p(\{\entry_k\}\,|\,\ratepars)$ as a piecewise
constant rate function---where the bin heights are the parameters---and if the
uncertainties on the catalog are negligible compared to the bin widths then
the maximum marginalized likelihood solution for \ratepars\ is a histogram of
the catalog entries (see \app{histogram} for the derivation).
The case of non-negligible uncertainties is described by \citet{hogge} using a
method similar to the one discussed here.

@@

\paragraph{Occurence rate parameterization \& priors}
In this paper, we will also model
